---
title: "Visualizing and summarizing numerical data"
format: 
  revealjs:
    auto-stretch: true
---

```{r}
#| include: false

library(tidyverse)
library(openintro)

ggplot2::theme_set(ggplot2::theme_gray(base_size = 16))
```

# Data

## Data: Lending Club

```{r}
loans <- loans_full_schema |>
  select(
    loan_amount, interest_rate, term, grade, 
    state, annual_income, homeownership, debt_to_income
  )
glimpse(loans)
```

# Histogram {.section-break}

## Histogram {.nostretch}

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram()
```

## Histograms and binwidth {.nostretch}

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 1000)
```

## Histograms and binwidth {.nostretch}

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 5000)
```

## Histograms and binwidth {.nostretch}

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 20000)
```

## Customizing histograms

```{r}
#| output-location: column
#| code-line-numbers: "|2|3-7"
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 5000) +
  labs(
    x = "Loan amount ($)",
    y = "Frequency",
    title = "Amounts of Lending Club loans"
  )
```

## Fill with a categorical variable

```{r}
#| output-location: column
#| code-line-numbers: "|5,10"
#| fig-asp: 0.8
ggplot(
  loans, 
  aes(
    x = loan_amount, 
    fill = homeownership
    )
  ) +
  geom_histogram(
    binwidth = 5000,
    alpha = 0.5
  ) +
  labs(
    x = "Loan amount ($)", y = "Frequency",
    title = "Amounts of Lending Club loans",
    fill = "Homeownership"
  ) +
  theme(legend.position = "bottom")
```

## Facet with a categorical variable

```{r}
#| output-location: column
#| code-line-numbers: "|14"
#| fig-asp: 1
ggplot(
  loans, 
  aes(
    x = loan_amount, 
    fill = homeownership
    )
  ) +
  geom_histogram(binwidth = 5000) +
  labs(
    x = "Loan amount ($)", y = "Frequency",
    title = "Amounts of Lending Club loans",
    fill = "Homeownership"
  ) +
  facet_wrap(~ homeownership, nrow = 3) +
  theme(legend.position = "bottom")
```

# Density plot

## Density plot {.nostretch}

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_density()
```

## Customizing density plots {.nostretch}

```{r}
#| code-line-numbers: "|3-7"
ggplot(loans, aes(x = loan_amount)) +
  geom_density() +
  labs(
    x = "Loan amount ($)", y = "Density",
    title = "Amounts of Lending Club loans"
  )
```

## Adding a categorical variable

```{r}
#| output-location: column
#| code-line-numbers: "|5,8"
ggplot(
  loans,
  aes(
    x = loan_amount, 
    fill = homeownership
    )
  ) +
  geom_density(alpha = 0.5) +
  labs(
    x = "Loan amount ($)", y = "Density",
    title = "Amounts of Lending Club loans", 
    fill = "Homeownership"
  )
```

# Box plot

## Box plot {.nostretch}

```{r}
ggplot(loans, aes(x = interest_rate)) +
  geom_boxplot()
```

## Box plot and outliers {.nostretch}

```{r}
ggplot(loans, aes(x = annual_income)) +
  geom_boxplot()
```

## Customizing box plots

```{r}
#| output-location: column
#| code-line-numbers: "|7-10"
ggplot(loans, aes(x = interest_rate)) +
  geom_boxplot() +
  labs(
    x = "Interest rate (%)", y = NULL,
    title = "Interest rates of Lending Club loans"
  ) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
```

## Adding a categorical variable

```{r}
#| output-location: column
#| code-line-numbers: "|3"
ggplot(
  loans, 
  aes(x = interest_rate, y = grade)) +
  geom_boxplot() +
  labs(
    x = "Interest rate (%)",
    y = "Grade",
    title = "Interest rates of Lending Club loans",
    subtitle = "by grade of loan"
  )
```
