---
title: "Diving deeper with Palmer Penguins (Complete)"
execute: 
  warning: false
fig-width: 6
fig-asp: 0.618
---

# Introduction

How do sizes of penguins vary across species, islands, and sexes?
What about other characteristics?

## Packages

We will use the **tidyverse** and **ggbeeswarm** packages for data wrangling and visualization and the **palmerpenguins** package for the data.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(palmerpenguins)
library(ggbeeswarm)
```

## Data

The dataset we will visualize is called `penguins`.
Let's `glimpse()` at it.

```{r}
#| label: glimpse
glimpse(penguins)
```

# Visualizing penguin weights

## Single variable

::: callout-note
Analyzing the a single variable is called **univariate** analysis.
:::

Create visualizations of the distribution of `weight`s of penguins.

### Exercise 1

Make a histogram.
Set an appropriate binwidth.

```{r}
#| label: histogram-weight
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 250)
```

### Exercise 2

Make a boxplot.

```{r}
#| label: boxplot-weight
#| fig-asp: 0.3
ggplot(penguins, aes(x = body_mass_g)) +
  geom_boxplot()
```

### Exercise 3

Based on these, determine if each of the following statements about the shape of the distribution is true or false.

-   The distribution of penguin weights in this sample is left skewed. **FALSE**
-   The distribution of penguin weights in this sample is unimodal. **Difficult to determine, FALSE**

```{r}
ggplot(penguins, aes(x = body_mass_g)) +
  geom_density()
```

## Two variables

::: callout-note
Analyzing the relationship between two variables is called **bivariate** analysis.
:::

Create visualizations of the distribution of `weight`s of penguins by `species`.

### Exercise 4

Make a single histogram.
Set an appropriate binwidth.

```{r}
#| label: histogram-weight-species
ggplot(penguins, aes(x = body_mass_g, fill = species)) +
  geom_histogram(binwidth = 250)
```

### Exercise 5

Use multiple histograms via faceting, one for each species.
Set an appropriate binwidth, add color as you see fit, and turn off legends if not needed.

```{r}
#| label: histogram-weight-species-facet
ggplot(penguins, aes(x = body_mass_g, fill = species)) +
  geom_histogram(binwidth = 250) +
  facet_wrap(~species, ncol = 1)
```

### Exercise 6

Use side-by-side box plots.
Add color as you see fit and turn off legends if not needed.

```{r}
#| label: boxplot-weight-species
ggplot(penguins, aes(x = body_mass_g, y = species, fill = species)) +
  geom_boxplot(show.legend = FALSE)
```

### Exercise 7

Use density plots.
Add color as you see fit.

```{r}
#| label: density-weight-species
ggplot(penguins, aes(x = body_mass_g, fill = species, color = species)) +
  geom_density(alpha = 0.5)
```

### Exercise 8

Use violin plots.
Add color as you see fit and turn off legends if not needed.

```{r}
#| label: violin-weight-species
ggplot(penguins, aes(x = body_mass_g, y = species, color = species, fill = species)) +
  geom_violin(alpha = 0.2, show.legend = FALSE)
```

### Exercise 9

Make a jittered scatter plot.
Add color as you see fit and turn off legends if not needed.

```{r}
#| label: jitter-weight-species
set.seed(1234)
ggplot(penguins, aes(x = species, y = body_mass_g)) +
  geom_jitter(alpha = 0.5)
```

### Exercise 10

Use beeswarm plots.
Add color as you see fit and turn off legends if not needed.

```{r}
#| label: beeswarm-weight-species
ggplot(penguins, aes(x = species, y = body_mass_g, color = species, shape = species)) +
  geom_beeswarm(show.legend = FALSE)
```

### Exercise 11

Use multiple geoms on a single plot.
Be deliberate about the order of plotting.
Change the theme and the color scale of the plot.
Finally, add informative labels.

```{r}
#| label: multi-geoms
ggplot(penguins, aes(x = body_mass_g, y = species, color = species)) +
  geom_boxplot() +
  geom_beeswarm()
```

## Multiple variables

::: callout-note
Analyzing the relationship between three or more variables is called **multivariate** analysis.
:::

### Exercise 12

Facet the plot you created in the previous exercise by `island`.
Adjust labels accordingly.

```{r}
#| label: multi-geoms-facet
ggplot(penguins, aes(x = body_mass_g, y = species, color = species)) +
  geom_boxplot(show.legend = FALSE) +
  geom_beeswarm(show.legend = FALSE) +
  facet_wrap(~island, ncol = 1) +
  labs(
    x = "Weight (grams)",
    y = "Species",
    title = "Weights of penguins across species and islands"
  )
```

Before you continue, let's turn off all warnings the code chunks generate and resize all figures.
We'll do this by editing the YAML.

# Visualizing other variables

### Exercise 13

Pick a single categorical variable from the data set and make a bar plot of its distribution.

```{r}
#| label: barplot
# add code here
```

### Exercise 14

Pick two categorical variables and make a visualization to visualize the relationship between the two variables.
Along with your code and output, provide an interpretation of the visualization.

Add interpretation here.

```{r}
#| label: cat-cat
# add code here
```

### Exercise 15

Make another plot that uses at least three variables.
At least one should be numeric and at least one categorical.
In 1-2 sentences, describe what the plot shows about the relationships between the variables you plotted.
Don't forget to label your code chunk.

Add interpretation here.

```{r}
# add code here
```
