---
title: "Diving deeper with Palmer Penguins"
---

# Introduction

How do sizes of penguins vary across species, islands, and sexes?
What about other characteristics?

## Packages

We will use the **tidyverse** packages for data wrangling and visualization and the **palmerpenguins** package for the data.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(palmerpenguins)
```

## Data

The dataset we will visualize is called `penguins`.
Let's `glimpse()` at it.

```{r}
#| label: glimpse
# add code here
```

# Visualizing penguin weights

## Single variable

::: callout-note
Analyzing the a single variable is called **univariate** analysis.
:::

Create visualizations of the distribution of `weight`s of penguins.

### Exercise 1

Make a histogram.
Set an appropriate binwidth.

```{r}
#| label: histogram-weight
# add code here
```

### Exercise 2

Make a boxplot.

```{r}
#| label: boxplot-weight
# add code here
```

### Exercise 3

Based on these, determine if each of the following statements about the shape of the distribution is true or false.

-   The distribution of penguin weights in this sample is left skewed.
-   The distribution of penguin weights in this sample is unimodal.

## Two variables

::: callout-note
Analyzing the relationship between two variables is called **bivariate** analysis.
:::

Create visualizations of the distribution of `weight`s of penguins by `species`.

### Exercise 4

Make a single histogram.
Set an appropriate binwidth.

```{r}
#| label: histogram-weight-species
# add code here
```

### Exercise 5

Use multiple histograms via faceting, one for each species.
Set an appropriate binwidth, add color as you see fit, and turn off legends if not needed.

```{r}
#| label: histogram-weight-species-facet
# add code here
```

### Exercise 6

Use side-by-side box plots.
Add color as you see fit and turn off legends if not needed.

```{r}
#| label: boxplot-weight-species
# add code here
```

### Exercise 7

Use density plots.
Add color as you see fit.

```{r}
#| label: density-weight-species
# add code here
```

### Exercise 8

Use violin plots.
Add color as you see fit and turn off legends if not needed.

```{r}
#| label: violin-weight-species
# add code here
```

### Exercise 9

Make a jittered scatter plot.
Add color as you see fit and turn off legends if not needed.

```{r}
#| label: jitter-weight-species
# add code here
```

### Exercise 10

Use beeswarm plots.
Add color as you see fit and turn off legends if not needed.

```{r}
#| label: beeswarm-weight-species
# add code here
```

### Exercise 11

Use multiple geoms on a single plot.
Be deliberate about the order of plotting.
Change the theme and the color scale of the plot.
Finally, add informative labels.

```{r}
#| label: multi-geoms
# add code here
```

## Multiple variables

::: callout-note
Analyzing the relationship between three or more variables is called **multivariate** analysis.
:::

### Exercise 12

Facet the plot you created in the previous exercise by `island`.
Adjust labels accordingly.

```{r}
#| label: multi-geoms-facet
# add code here
```

Before you continue, let's turn off all warnings the code chunks generate and resize all figures.
We'll do this by editing the YAML.

# Visualizing other variables

### Exercise 13

Pick a single categorical variable from the data set and make a bar plot of its distribution.

```{r}
#| label: barplot
# add code here
```

### Exercise 14

Pick two categorical variables and make a visualization to visualize the relationship between the two variables.
Along with your code and output, provide an interpretation of the visualization.

Add interpretation here.

```{r}
#| label: cat-cat
# add code here
```

### Exercise 15

Make another plot that uses at least three variables.
At least one should be numeric and at least one categorical.
In 1-2 sentences, describe what the plot shows about the relationships between the variables you plotted.
Don't forget to label your code chunk.

Add interpretation here.

```{r}
# add code here
```
