---
title: "Flights and pipes"
subtitle: "End of video"
---

# Introduction

Have you ever had a flight delayed?
How common is that?
Which airlines are delayed more than others?

## Packages

We will use the **tidyverse** packages for data wrangling and visualization and the **nycflights13** package for the data.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(nycflights13)
```

## Data

The dataset we will explore is called `flights`.
Let's `glimpse()` at it.

```{r}
#| label: glimpse-flights
glimpse(flights)
```

# Exploring flights

## Exercise 1

Fill in the blanks:

The `flights` data frame has `r nrow(flights)` rows.
Each row represents a flight out of one of three NYC airports in 2013.

## Exercise 2

What are the names of the variables in `flights`.

```{r}
#| label: column-names
names(flights)
```

## Exercise 3 - `select()`

-   Make a data frame that only contains the variables `dep_delay` and `arr_delay`.

```{r}
#| label: select-vars
flights |>
  select(dep_delay, arr_delay)
```

-   Make a data frame that keeps every variable except `dep_delay`.

```{r}
#| label: exclude-vars
flights |>
  select(!dep_delay)
```

-   Make a data frame that includes all variables between `year` through `dep_delay` (inclusive). These are all variables that provide information about the departure of each flight.

```{r}
#| label: include-range
flights |>
  select(year:dep_delay)
```

-   Use the `select` helper `contains()` to make a data frame that includes the variables associated with the arrival, i.e., contains the string `"arr_"` in the name.

```{r}
#| label: arr-vars
flights |>
  select(contains("arr_"))
```

## Exercise 4 - `slice()`

-   Display the first five rows of the `flights` data frame.

```{r}
#| label: slice
flights |>
  slice(1:5)
```

-   Display the last two rows of the `flights` data frame.

```{r}
#| label: last-two
flights |>
  slice_tail(n = 2)
```

## Exercise 5 - `arrange()`

-   Let's arrange the data by departure delay, so the flights with the shortest departure delays will be at the top of the data frame.

```{r}
#| label: arrange-delays
flights |>
  arrange(dep_delay)
```

-   What does it mean for the `dep_delay` to have a negative value?

Flight took off 43 minutes earlier than it was scheduled.

-   Arrange the data by descending departure delay, so the flights with the longest departure delays will be at the top.

```{r}
#| label: arrange-delays-desc
flights |>
  arrange(desc(dep_delay))
```

-   Create a data frame that only includes the plane tail number (`tailnum`), carrier (`carrier`), and departure delay for the flight with the longest departure delay. What is the plane tail number (`tailnum`) for this flight?

```{r}
#| label: longest-delay
flights |>
  arrange(desc(dep_delay)) |>
  slice_head(n = 1) |>
  select(tailnum, carrier, dep_delay)
```

## Exercise 6 - `filter()`

-   Filter for all rows where the destination airport is RDU.

```{r}
#| label: rdu
flights |>
  filter(dest == "RDU")
```

-   Filter for all rows where the destination airport is RDU and the arrival delay is less than 0.

```{r}
#| label: rdu-ontime
flights |>
  filter(dest == "RDU" & arr_delay < 0)
```

-   Describe what the code is doing in words.

Start with the flights data frame, and then filter for rows where destination is RDU or GSO and arrival delay or departure delay is less than 0.

```{r}
#| label: nc-early
flights |>
  filter(
    dest %in% c("RDU", "GSO"),
    arr_delay < 0 | dep_delay < 0
  )
```

**Hint:** Logical operators in R:

| operator      | definition                                               |
|:--------------|:---------------------------------------------------------|
| `<`           | is less than?                                            |
| `<=`          | is less than or equal to?                                |
| `>`           | is greater than?                                         |
| `>=`          | is greater than or equal to?                             |
| `==`          | is exactly equal to?                                     |
| `!=`          | is not equal to?                                         |
| `x & y`       | is x AND y?                                              |
| `x | y`       | is x OR y?                                               |
| `is.na(x)`    | is x NA?                                                 |
| `!is.na(x)`   | is x not NA?                                             |
| `x %in% y`    | is x in y?                                               |
| `!(x %in% y)` | is x not in y?                                           |
| `!x`          | is not x? (only makes sense if `x` is `TRUE` or `FALSE`) |

## Exercise 7 - `count()`

-   Create a frequency table of the destination locations for flights from New York.

```{r}
#| label: count-dest
flights |>
  count(dest, sort = TRUE)
```

-   In which month was there the fewest number of flights? How many flights were there in that month?

```{r}
#| label: count-month
flights |>
  count(month) |>
  arrange(n)
```

-   On which date (month + day) was there the largest number of flights? How many flights were there on that day?

```{r}
#| label: count-date
flights |>
  count(month, day, sort = TRUE)
```

## Exercise 8 - `mutate()`

-   Convert `air_time` (minutes in the air) to hours and then create a new variable, `mph`, the miles per hour of the flight.

```{r}
#| label: calculate-mph
flights |>
  mutate(
    air_time_hr = air_time / 60,
    mph = distance / air_time_hr
  )
```

-   First, count the number of flights each month, and then calculate the proportion of flights in each month. What proportion of flights take place in July?

```{r}
#| label: months-prop
flights |>
  count(month) |>
  mutate(prop = n / sum(n)) |>
  filter(month == 7)
```

## Exercise 9 - `summarize()`

-   Find mean arrival delay for all flights.

```{r}
#| label: find-arrival-delay
flights |>
  summarize(mean_arr_delay = mean(arr_delay, na.rm = TRUE))
```

## Exercise 10 - `group_by()`

-   Find mean arrival delay for for each month.

```{r}
#| label: mean-arr-delay-month
flights |>
  group_by(month) |>
  summarize(mean_arr_delay = mean(arr_delay, na.rm = TRUE))
```

-   What is the median departure delay for each airports around NYC (`origin`)? Which airport has the shortest median departure delay?

```{r}
#| label: median-dep-dely-origin
flights |>
  group_by(origin) |>
  summarize(median_dep_delay = median(dep_delay, na.rm = TRUE))
```
