---
title: "Starwars characters"
editor_options: 
  chunk_output_type: console
---

# Introduction

You might not be a Star Wars fan, but you've probably heard of the movie franchise!

## Packages

We will use the **tidyverse** packages for data wrangling and for the data.

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
```

## Data

The dataset we will explore is called `starwars`.

# Star Wars characters

## Exercise 1

Glimpse at the starwars data frame.

```{r}
#| label: glimpse-starwars

glimpse(starwars)
```

## Exercise 2

Modify the following plot to change the color of the points by gender.

```{r}
#| label: scatterplot-gender

ggplot(starwars, aes(x = height, y = mass, color = gender)) +
  geom_point()
```

## Exercise 3

Modify the following plot to change the color of all points to `"pink"`.

```{r}
#| label: scatterplot-pink

ggplot(starwars, aes(x = height, y = mass)) +
  geom_point(color = "pink")
```

## Exercise 4

Add labels for title, x and y axes, and color of points.

```{r}
#| label: scatterplot-labels

ggplot(starwars, aes(x = height, y = mass, color = gender, shape = gender)) +
  geom_point() +
  labs(
    title = "Weights and heights of Star Wars characters",
    x = "Height (cm)",
    y = "Weight (kg)",
    color = "Gender",
    shape = "Gender"
  )
```

## Exercise 5

Pick a single categorical variable from the data set and make a bar plot of its distribution.
Describe what you see in the plot.

Add description here.

```{r}
#| label: viz-cat

ggplot(starwars, aes(y = hair_color)) +
  geom_bar()
```

## Exercise 6

Pick two categorical variables and make a visualization to visualize the relationship between the these variables.
Along with your code and output, provide an interpretation of the visualization.

Add interpretation here.

```{r}
#| label: viz-cat-cat

ggplot(starwars, aes(x = sex, fill = gender)) +
  geom_bar()

ggplot(starwars, aes(x = gender, fill = sex)) +
  geom_bar()
```

## Exercise 7

Pick three categorical variables and make a visualization to visualize the relationship between the these variables.
Along with your code and output, provide an interpretation of the visualization.

Add interpretation here.

```{r}
#| label: viz-cat-cat-cat

ggplot(starwars, aes(x = gender, fill = sex)) +
  geom_bar() +
  facet_wrap(~hair_color)
```
