---
title: "Visualizing and summarizing relationships"
format: 
  revealjs:
    auto-stretch: true
---

```{r}
#| include: false
ggplot2::theme_set(ggplot2::theme_gray(base_size = 16))
```

# Setup

## Packages

```{r}
#| label: packages
#| message: false
library(tidyverse) # for data wrangling and visualization
library(openintro) # for data
library(scales)    # for pretty axis labels
```

## Data: Houses in Duke Forest

```{r}
glimpse(duke_forest)
```

## Price vs. area {.nostretch}

```{r}
#| echo: false
#| fig-width: 12
ggplot(duke_forest, aes(x = area, y = price, color = cooling, shape = cooling)) +
  geom_point(alpha = 0.7, size = 4) +
  scale_x_continuous(labels = label_number(big.mark = ",")) +
  scale_y_continuous(labels = label_dollar()) +
  scale_color_manual(values = c("central" = "darkblue", "other" = "lightsalmon3")) +
  labs(
    title = "Houses in Duke Forest",
    subtitle = "Durham, NC",
    color = "Cooling", shape = "Cooling",
    x = "Area (square feet)",
    y = "Price (USD)",
    caption = "Source: Zillow, November 2020"
  ) +
  theme_minimal(base_size = 26) +
  theme(legend.position = c(0.15, 0.8))
```

# Relationship between two numerical variables

## Scatterplot

```{r}
#| label: area-price-scatter
ggplot(duke_forest, aes(x = area, y = price)) +
  geom_point()
```

## Correlation coefficient

```{r}
duke_forest |>
  summarize(r = cor(area, price))
```

## Characteristics of a relationship

*between two numerical variables*

::: columns
::: {.column .fragment width="30%"}
::: incremental
-   **Direction:** Positive

-   **Form:** Linear

-   **Strength:** Moderately strong
:::
:::

::: {.column width="40%"}
```{r}
#| ref.label: area-price-scatter
#| fig-width: 5
#| echo: false
```
:::
:::

# Relationship between a numerical and a categorical variable

## Distribution of house prices

```{r}
ggplot(duke_forest, aes(x = price)) +
  geom_histogram(binwidth = 100000)
```

## Filled histograms

```{r}
ggplot(duke_forest, aes(x = price, fill = cooling)) +
  geom_histogram(binwidth = 100000)
```
  
## Side-by-side box plots

```{r}
ggplot(duke_forest, aes(x = price, y = cooling)) +
  geom_boxplot()
```

## Filled density plots

```{r}
ggplot(
  duke_forest, 
  aes(x = price, color = cooling, fill = cooling)
  ) +
  geom_density(alpha = 0.5)
```

## Violin plots

```{r}
ggplot(duke_forest, aes(x = price, y = cooling)) +
  geom_violin()
```

## Grouped summaries

```{r}
duke_forest |>
  summarize(median_price = median(price))
```

. . .

```{r}
#| code-line-numbers: "|2"
duke_forest |>
  group_by(cooling) |>
  summarize(
    median_price = median(price), 
  )
```

## Multiple summary statistics

```{r}
#| code-line-numbers: "|4-7"
duke_forest |>
  group_by(cooling) |>
  summarize(
    min_price = min(price),
    median_price = median(price), 
    iqr_price = IQR(price),
    max_price = max(price)
  )
```

# Relationship between two categorical variables

## Year built

::: task
Create a new variable called `year_built_cat` that takes on the value `"1970 or later"` if the house is built in 1970 or later, and `"1969 or earlier"` otherwise.
:::

. . .

```{r}
#| code-line-numbers: "|2|3-7|4|5|6|1"
duke_forest <- duke_forest |>
  mutate(
    year_built_cat = if_else(
      year_built >= 1970, 
      "1970 or later", 
      "1969 or earlier"
    )
   )
```

## Year built

```{r}
duke_forest |>
  select(contains("year_built"))
```

## Stacked bar plot -- frequencites

```{r}
ggplot(duke_forest, aes(x = year_built_cat, fill = cooling)) +
  geom_bar()
```

## Stacked bar plot -- proportions

```{r}
ggplot(duke_forest, aes(x = year_built_cat, fill = cooling)) +
  geom_bar(position = "fill")
```

## Frequencies

```{r}
duke_forest |>
  count(year_built_cat, cooling)
```

## Proportions

```{r}
duke_forest |>
  count(year_built_cat, cooling) |>
  group_by(year_built_cat) |>
  mutate(prop = n / sum(n))
```

# Relationships between many variables

## Price vs. area and cooling {.smaller}

```{r}
#| output-location: column
#| code-line-numbers: "|1,4|2|3|5"
ggplot(
  duke_forest, 
  aes(x = area, y = price, color = cooling, shape = cooling)
  ) +
  geom_point()
```

## Price vs. area and cooling {.smaller}

```{r}
#| output-location: column
#| code-line-numbers: "5"
ggplot(
  duke_forest, 
  aes(x = area, y = price, color = cooling, shape = cooling)
  ) +
  geom_point(alpha = 0.7, size = 4)
```

## Price vs. area and cooling {.smaller}

```{r}
#| output-location: column
#| code-line-numbers: "6-13"
ggplot(
  duke_forest, 
  aes(x = area, y = price, color = cooling, shape = cooling)
  ) +
  geom_point(alpha = 0.7, size = 4) +
  labs(
    title = "Houses in Duke Forest",
    subtitle = "Durham, NC",
    color = "Cooling", shape = "Cooling",
    x = "Area (square feet)",
    y = "Price (USD)",
    caption = "Source: Zillow, November 2020"
  )
```

## Price vs. area and cooling {.smaller}

```{r}
#| output-location: column
#| code-line-numbers: "6"
ggplot(
  duke_forest, 
  aes(x = area, y = price, color = cooling, shape = cooling)
  ) +
  geom_point(alpha = 0.7, size = 4) +
  scale_color_manual(values = c("central" = "darkblue", "other" = "lightsalmon3")) +
  labs(
    title = "Houses in Duke Forest",
    subtitle = "Durham, NC",
    color = "Cooling", shape = "Cooling",
    x = "Area (square feet)",
    y = "Price (USD)",
    caption = "Source: Zillow, November 2020"
  )
```

## Price vs. area and cooling {.smaller}

```{r}
#| output-location: column
#| code-line-numbers: "6"
ggplot(
  duke_forest, 
  aes(x = area, y = price, color = cooling, shape = cooling)
  ) +
  geom_point(alpha = 0.7, size = 4) +
  scale_x_continuous(labels = label_number(big.mark = ",")) +
  scale_color_manual(values = c("central" = "darkblue", "other" = "lightsalmon3")) +
  labs(
    title = "Houses in Duke Forest",
    subtitle = "Durham, NC",
    color = "Cooling", shape = "Cooling",
    x = "Area (square feet)",
    y = "Price (USD)",
    caption = "Source: Zillow, November 2020"
  )
```

## Price vs. area and cooling {.smaller}

```{r}
#| output-location: column
#| code-line-numbers: "7"
ggplot(
  duke_forest, 
  aes(x = area, y = price, color = cooling, shape = cooling)
  ) +
  geom_point(alpha = 0.7, size = 4) +
  scale_x_continuous(labels = label_number(big.mark = ",")) +
  scale_y_continuous(labels = label_dollar()) +
  scale_color_manual(values = c("central" = "darkblue", "other" = "lightsalmon3")) +
  labs(
    title = "Houses in Duke Forest",
    subtitle = "Durham, NC",
    color = "Cooling", shape = "Cooling",
    x = "Area (square feet)",
    y = "Price (USD)",
    caption = "Source: Zillow, November 2020"
  )
```

## Price vs. area and cooling {.smaller}

```{r}
#| output-location: column
#| code-line-numbers: "17"
ggplot(
  duke_forest, 
  aes(x = area, y = price, color = cooling, shape = cooling)
  ) +
  geom_point(alpha = 0.7, size = 4) +
  scale_x_continuous(labels = label_number(big.mark = ",")) +
  scale_y_continuous(labels = label_dollar()) +
  scale_color_manual(values = c("central" = "darkblue", "other" = "lightsalmon3")) +
  labs(
    title = "Houses in Duke Forest",
    subtitle = "Durham, NC",
    color = "Cooling", shape = "Cooling",
    x = "Area (square feet)",
    y = "Price (USD)",
    caption = "Source: Zillow, November 2020"
  ) +
  theme_minimal()
```

## Price vs. area and cooling {.smaller}

```{r}
#| output-location: column
#| code-line-numbers: "18"
ggplot(
  duke_forest, 
  aes(x = area, y = price, color = cooling, shape = cooling)
  ) +
  geom_point(alpha = 0.7, size = 4) +
  scale_x_continuous(labels = label_number(big.mark = ",")) +
  scale_y_continuous(labels = label_dollar()) +
  scale_color_manual(values = c("central" = "darkblue", "other" = "lightsalmon3")) +
  labs(
    title = "Houses in Duke Forest",
    subtitle = "Durham, NC",
    color = "Cooling", shape = "Cooling",
    x = "Area (square feet)",
    y = "Price (USD)",
    caption = "Source: Zillow, November 2020"
  ) +
  theme_minimal() +
  theme(legend.position = c(0.15, 0.8))
```

## Price vs. area and cooling

```{r}
duke_forest |>
  group_by(cooling) |>
  summarize(r = cor(price, area))
```

## Price vs. area and cooling  {.smaller}

```{r}
#| output-location: column
#| code-line-numbers: "18"
#| fig-asp: 1
ggplot(
  duke_forest, 
  aes(x = area, y = price, color = cooling, shape = cooling)
  ) +
  geom_point(alpha = 0.7, size = 4) +
  scale_x_continuous(labels = label_number(big.mark = ",")) +
  scale_y_continuous(labels = label_dollar()) +
  scale_color_manual(values = c("central" = "darkblue", "other" = "lightsalmon3")) +
  labs(
    title = "Houses in Duke Forest",
    subtitle = "Durham, NC",
    color = "Cooling", shape = "Cooling",
    x = "Area (square feet)",
    y = "Price (USD)",
    caption = "Source: Zillow, November 2020"
  ) +
  theme_minimal() +
  facet_wrap(~cooling, ncol = 1)
```

## Price vs. area and cooling {.smaller}

```{r}
#| output-location: column
#| code-line-numbers: "5,10"
#| fig-asp: 1
ggplot(
  duke_forest, 
  aes(x = area, y = price, color = cooling, shape = cooling)
  ) +
  geom_point(alpha = 0.7, size = 4, show.legend = FALSE) +
  scale_x_continuous(labels = label_number(big.mark = ",")) +
  scale_y_continuous(labels = label_dollar()) +
  scale_color_manual(values = c("central" = "darkblue", "other" = "lightsalmon3")) +
  labs(
    title = "Houses in Duke Forest, by cooling type",
    subtitle = "Durham, NC",
    color = "Cooling", shape = "Cooling",
    x = "Area (square feet)",
    y = "Price (USD)",
    caption = "Source: Zillow, November 2020"
  ) +
  theme_minimal() +
  facet_wrap(~cooling, ncol = 1)
```
