---
title: "First look at Palmer Penguins"
---

# Introduction

How do bill sizes of penguins vary across species?
And what happens if we don't take species into consideration?

## Packages

We will use the **tidyverse** packages for data wrangling and visualization and the **palmerpenguins** package for the data.

```{r}
#| label: load-packages
library(tidyverse)
library(palmerpenguins)
```

## Data

The dataset we will visualize is called `penguins`.
Let's `glimpse()` at it.

```{r}
#| label: glimpse
# add code here
```

# Penguin bills

1.  Create a scatterplot of `bill_depth` vs. `bill_length`s of `penguin`s. Overlay a line of best fit. Describe the relationship between the two variables.

```{r}
#| label: bill-depth-width
# add code here
```

Add interpretation here.

2.  Create a scatterplot of `bill_depth` vs. `bill_length`s of `penguin`s, colored by `species`. Overlay a line of best fit. Describe the relationship between the two variables for each of the species.

```{r}
#| label: bill-depth-width-species
# add code here
```

Add interpretation here.

3.  Reflect on the seemingly contradictory findings from the two visualizations you've created. Which one do you believe more, and why?
