---
title: "First look at Palmer Penguins (Complete)"
---

# Introduction

How do bill sizes of penguins vary across species?
And what happens if we don't take species into consideration?

## Packages

We will use the **tidyverse** packages for data wrangling and visualization and the **palmerpenguins** package for the data.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(palmerpenguins)
```

## Data

The dataset we will visualize is called `penguins`.
Let's `glimpse()` at it.

```{r}
#| label: glimpse
glimpse(penguins)
```

# Penguin bills

1.  Create a scatterplot of `bill_depth_mm` vs. `bill_length_mm`s of `penguin`s. Overlay a line of best fit. Describe the relationship between the two variables.

```{r}
#| label: bill-depth-width
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

There is a negative relationship between bill depth and bill length.
The relationship is somewhat weak and linear.

2.  Create a scatterplot of `bill_depth_mm` vs. `bill_length_mm`s of `penguin`s, colored by `species`. Overlay a line of best fit. Describe the relationship between the two variables for each of the species.

```{r}
#| label: bill-depth-width-species
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

There is a positive relationship between bill depth and bill length for each of the three species.
The relationships are moderately strong and linear.

3.  Reflect on the seemingly contradictory findings from the two visualizations you've created. Which one do you believe more, and why?

I believe the second one more, we should be considering species when analyzing the relationship between bill depth and bill length for penguins.
